---
title: "Michael Havadjias PSTAT 131 HW1"
output: html_document
date: "2022-09-30"
---

**Question 1:
Define supervised and unsupervised learning. What are the difference(s) between them?**


  In supervised learning for each observation of the predictor measurement(s) xi, i = 1,...,n there is an associated response measurement yi. In unsupervised learning for every observation i = 1,...,n, we observea vector of measurements xi but no associated response yi. (From book page # 26)

**Question 2:
Explain the difference between a regression model and a classification model, specifically in the context of machine learning.**

 We tend to refer to problems with a quantitative response as regression problems, while those involving a qualitative response are often referred to as classification problems. (pg #28)
 
**Question 3:
Name two commonly used metrics for regression ML problems. Name two commonly used metrics for classification ML problems.**

Regression : Training MSE and Test MSE

Classification: Training Error Rate and Test Error Rate.
 
**Question 4:
As discussed, statistical models can be used for different purposes. These purposes can generally be classified into the following three categories. Provide a brief description of each.**

**Descriptive models:**
  
  Used to emphasize trends in data visually ( from lecture)

**Inferential models:**
  
  Aim to test theories and state the relationship between outcome and predictors (from lecture)
  
**Predictive models:**
  
  Are not focused on hypothesis tests and aim to predict Y with minimum reducible error (from lecture)
 
 
**Question 5:
Predictive models are frequently used in machine learning, and they can usually be described as either mechanistic or empirically-driven. Answer the following questions.**

**Define mechanistic. Define empirically-driven. How do these model types differ? How are they similar?**
  
  A mechanistic model is a model that uses theory to make a prediction whereas an empirically driven model relies on observations rather than theory. Although both models can become overfitted with too many observations empirically driven models generally require more observations and are more flexible by default. In addition to that empirically driven models do not make assumptions about f whereas mechanistic models assume a parametric form for f. 
  
**In general, is a mechanistic or empirically-driven model easier to understand? Explain your choice.**

An empirically driven model is easier to understand as you can reliably predict an outcome following an occasion based on observation whereas in a mechanistic model you need to rely on theory to predict the outcome.

**Describe how the bias-variance tradeoff is related to the use of mechanistic or empirically-driven models.**
  
  In general, more flexible statistical methods have higher variance and result in less bias. As a general rule, as we use more flexible methods, the variance will increase and the bias will decrease (pg #35 ). Since empirically driven models are more flexible by default we would expect them to generally have higher variance and less bias than mechanistic models.
  
**Question 6:
A political candidate’s campaign has collected some detailed voter history data from their constituents. The campaign is interested in two questions:**

**Given a voter’s profile/data, how likely is it that they will vote in favor of the candidate?**
  
  Predictive as it relies on the voters profile and data to predict their preference.


**How would a voter’s likelihood of support for the candidate change if they had personal contact with the candidate?**
 
 Inferential as it relies on a hypothesis to draw a conclusion about the population. The hypothesis being whether personal contact with a candidate will change a voters likelihood of support.
 
**Classify each question as either predictive or inferential. Explain your reasoning for each.**


 <h1>Exploratory Data Analysis1</h1>#
 
**Exercise 1:
We are interested in highway miles per gallon, or the hwy variable. Create a histogram of this variable. Describe what you see/learn.**

```{r}
library(tidyverse)
library(ggplot2)
hist(mpg$hwy,xlab = "Highway Miles per Gallon")
```

The histogram reveals that most cars make 25 to 30 miles per gallon on the highway with a frequency of approximately 80 cars. The second most occurring interval is the 15 to 20 miles per gallon range with approximately 70 cars.The 20 to 25 mpg has approximately 40 occurrences making it the 3rd most occurring interval. The 10-15 and 30-35 mpg intervals are very similar with about 18 cars in each category achieving those highway mpg values.  The frequency of cars making more than 35 miles to the gallon decreases dramatically with the ranges 35-40 and 40-45 hwy having approximately 3 cars each. The above histogram reveals that most automakers aim to achieve 15-30 highway miles per gallon. 


**Exercise 2:
Create a scatterplot. Put hwy on the x-axis and cty on the y-axis. Describe what you notice. Is there a relationship between hwy and cty? What does this mean?**

```{r}
plot(mpg$hwy,mpg$cty,xlab = "Highway Miles per galon",ylab = "City Miles Per Gallon")
```


 Highway miles per gallon and City miles per gallon are positively correlated, that is cars that have high hwy values also have high cty values. This indicates that as fuel efficiency goes up in terms of highway miles there is a corresponding increase in cty fuel efficiency. 



**Exercise 3:
Make a bar plot of manufacturer. Flip it so that the manufacturers are on the y-axis. Order the bars by height. Which manufacturer produced the most cars? Which produced the least?**

```{r}
ggplot(data = mpg, aes(x = manufacturer)) +
  geom_bar() +
  coord_flip()

```


Dodge produced the most cars and Lincoln produced the least.

**Exercise 4:
Make a box plot of hwy, grouped by cyl. Do you see a pattern? If so, what?**

```{r}
boxplot(hwy~cyl, data = mpg)
```

 As the number of cylinders increases fuel efficiency generally goes down. It also appears that vehicles with 6 cylinders have the highest variability in fuel efficiency. 5 cylinder vehicles appear to have the least variability and are closest to 4 cylinder vehicles in terms of fuel efficiency. 
 
 
 
**Exercise 5:
Use the corrplot package to make a lower triangle correlation matrix of the mpg dataset. (Hint: You can find information on the package here.)**

**Which variables are positively or negatively correlated with which others? Do these relationships make sense to you? Are there any that surprise you?**

```{r} 
library(corrplot)
mpg_data <- mpg[,c(3,4,5,8,9)] #select numeric columns from mpg data
cor_mpg_data<-cor(mpg_data) # computes correlation, variance and covariance
corrplot(corr = cor_mpg_data,method = "circle",type = "lower")
```

displ - year : Positive (weak)

displ - cyl : Strong positive

displ - cty : Negative 

displ - hwy : Negative

year - cyl : Positive (weak)

year - cty : Negative (almost no correlation)

year - hwy : Positive (almost no correlation)

cyl - cty : Negative (strong)

cyl - hwy : Negative (strong)

cty - hwy : Positive (strong)


Most relationships make sense as they are correlated in a way that makes sense. That is, displacement, number of cylinders and highway and city miles per gallon behave in a sensible manner in relation to one another. (eg displacement increases cty and hwy efficiency decreases or number of cylinders increases cty and hwy decrease.). What is a little surprising is the weak correlation between year and cyl as I would expect no correlation or a negative one as smaller engines are becoming more common as years go by.
















